\blue{...Numerical investigation of viscous flow instabilities through heterogeneous porous media using adaptive and unstructured meshes under different mobility scenarios...}

### 2017-07-31 Papers/ViscousInstabilities/Manuscript_ViscousInstabilities_KC.tex
updates

\medskip
As already mentioned in section \ref{section:ViscousInstabilities}, \textit{viscous fingering} is characterized by the formation of a characteristic pattern, called viscous fingers. The geometrical evolution of these patterns provides good indications on capturing and tracking any discontinuities or instabilities in the flow front. 
Figure \ref{fig:3DChannel_PermMesh} describes the permeability pattern as this is implied in a $3D$ domain. The surface on the left hand side of the domain is the area where fluid $1$ is injected with a velovity $ u=2.5 cm/sec $. In the areas around the domain a no-slip condition is implied (\ie insulation, no exchange of material or Neumann boundary condition for mass transfer) while at the right hand side of the domain is the outlet. The domain is constructed in such a way so as to be able to focus and capture the preferential path of flow as this can be seen from figures \ref{fig:3DChannel_planes}-\ref{fig:3DChannel_Isosurf}.

@@

\medskip
In previous sections \ref{section:results:setup}-\ref{section:results_hete_fix_adapt}, a $2$D domain was used to describe one or more finger formations. These flow instabilities were triggerded because of the permeability differences in the domain. Using the $3D$ domain, the difference between a flow instability and the preferential flow path that may include smaller fingers is more obvious. Figures \ref{fig:3DChannel_sat}(a)-(f) give a description of the preferable flow path in various time-steps alongside with the progression of the saturation front that now becomes more obvious. Based on the boundary conditions that implemented, the preferencial path of the flow can been seen. Heterogeneity in the reservoir controls the spatial distribution of porosity and permeability. The permeability depends on the number, size and connectivity of the pores. An on purpose implementation of significant high permeability zones (compared with the lower ones) also know as higher permeability channels, allows our test-cases to give a good approximation of the so called "thief zones".

\medskip 
The saturation front progress and the idea of preferential flow path is followed by a more detail representation as this can be found in figure \ref{fig:3DChannel_Isosurf}. For the same time-steps an inside view on the domain provides with a more detailed description of not only the flow path itself as well as the minor fingers that may be created. At this point resolution plays an important role on capturing these patterns in detail and is also a limititation that needs to be overcome.

\medskip
A quick review to the mechanisms of viscous fingering (\ie shielding, spreading and tip-splitting, \citet{homsy_1987}) is required. In all the $2$D previous cases (sections \ref{section:results:setup}-\ref{section:results_hete_fix_adapt}), as surface tension becomes weak, the front of the steady finger is susceptible to a viscous-fingering instability. Through \textit{shielding} mechanism, one of these fingers will eventually outgrow the other and, will then spread (\textit{spreading}) to occupy the appropriate width of the domain ahead of it. Following this process, it reaches a width that is again unstable and the pattern repeats. So far shielding and spreading were dominating the flow but using the current $3$D domain and because of the lack of high resolution, the conditions and the mechanics of tip-splitting along a front can be qualitatively understood.

>>>>

\medskip
As already mentioned in section \ref{section:ViscousInstabilities}, \textit{viscous fingering} is characterized by the formation of flow patterns, called viscous fingers. The geometrical evolution of these patterns provides good indications on capturing and tracking any discontinuities or instabilities in the flow front. Figure \ref{fig:3DChannel_PermMesh} describes the permeability pattern as this is implied in a $3D$ domain. The surface on the left hand side of the domain is the area where fluid $1$ is injected with a velovity $ u=2.5 cm/sec $. No-slip conditions are implied around the domain (\ie insulation, no exchange of material or Neumann boundary condition for mass transfer) while at the right hand side of the domain is the outlet. The control volume is constructed in such a way so as to be able to focus and capture the preferential path of flow as this can be seen from figures \ref{fig:3DChannel_planes}-\ref{fig:3DChannel_Isosurf}.

@@

\medskip
In previous sections \ref{section:results:setup}-\ref{section:results_hete_fix_adapt}, a $2$D domain was used to describe one or more finger formations. These flow instabilities were triggerded because of the permeability differences in the domain. Using the $3D$ domain, the difference between a flow instability and the preferential flow path that may include smaller fingers is more obvious. Figures \ref{fig:3DChannel_sat}(a)-(f) give a description of the preferable flow path in various time-steps alongside with the progression of the saturation front that now becomes more obvious. Based on the boundary conditions that implemented, the preferencial path of the flow can been seen. Heterogeneity in the reservoir controls the spatial distribution of porosity and permeability. The permeability depends on the number, size and connectivity of the pores. An on purpose implementation of significant high permeability zones (compared with the lower ones) also know as higher permeability channels, allows the domain to give a good approximation of the so called "thief zones".

\medskip 
The saturation front progress and the idea of preferential flow path is followed by a more detail representation as this can be found in figure \ref{fig:3DChannel_Isosurf}. For the same time-steps an inner view on the domain provides with a more detailed description of not only the flow path itself as well as the minor fingers that may be created. At this point resolution plays an important role on capturing these patterns in detail and is also a limititation that needs to be overcome.

\medskip
A quick review to the mechanisms of viscous fingering (\ie shielding, spreading and tip-splitting, \citet{homsy_1987}) is required. In all the $2$D previous cases (sections \ref{section:results:setup}-\ref{section:results_hete_fix_adapt}), as interfacial surface tension (IFT) becomes weak, the front of the steady finger is susceptible to a viscous-fingering instability. Through \textit{shielding} mechanism, one of these fingers will eventually outgrow the other and, will then spread (\textit{spreading}) to occupy the appropriate width of the domain ahead of it. Following this process, it reaches a width that is again unstable and the pattern repeats. So far shielding and spreading were dominating the flow but, using the current $3$D domain, the conditions and the mechanics of tip-splitting along a front can be qualitatively understood.
